Convolutional Neural Networks on Graphs\\ with Fast Localized Spectral Filtering

### 2016-10-29 nips2016/review/rebuttal_michael.txt

Todo
* compare with GNN
* read recent spatial approaches
* constructed graph on MNIST
* run on RCV1
* review "?" in text

Acceptance
* oral: ~2%
* spotlight: ~5%
* poster: ~22%

Private

It appears to us that we got good reviews from reviewers who consider themselves
experts or are confident. Less confident reviewers seem to have not understood
the paper, thus its potential. E.g. they ask questions whose answers are in the
paper already.

General

We thank all the reviewers for reading the paper and providing valuable
feedback. The specific comments of each reviewer are addressed below.

We already cleaned up and edited the paper in the hope to make it more
accessible. We'll take into account the feedbacks for the final revision.

We plan to publicly release the code once the paper is published so that people
can reproduce our results, compare with our model or apply it to their problems.

Note that very recent works [post-submission refs?] (posted on arXiv after the
submission deadline) shows the growing interest for a generalization of CNNs to
arbitrary graphs. While both these approaches are spatial, we look forward to
compare with them.

Reviewer 1: comparison with GNN.

GNN does embed the graph into an high-dimensional space.
Limited to combinatorial Laplacian, no normalized?.
Classify graphs (graph-focused) or nodes (node-focused).
Thanks for pointing this up. We'll add this discussion to the final version of
the paper.

First, the goal of GNNs is to classify graphs or nodes, while the proposed
method classifies signals on graphs. Then, each approach could be twisted and we
could treat a graph signal as labels, but then there'll be no pooling.
Our method may be used to classify nodes, by treating a graph signal as a set of
labels, but we don't see how GNNs could be used to classify graph signals.

Note suited for multiple signals on graph. But for features who leave on nodes.

GNNs is designed as a spatial method.

Second, GNNs need to perform as many backpropagation operations as there are
edges and nodes in the graph per learning iteration. Complexity of

Graph neural networks (GNNs) (Scarselli et al., 2009) are a recurrent neural
network architecture defined on graphs. GNNs apply recurrent neural networks for
walks on the graph structure, propagating node representations until a fixed
point is reached. The resulting node representations are then used as features
in classification and regression problems. GNNs support only discrete labels and
perform as many backpropagation operations as there are edges and nodes in the
graph per learning iteration. Gated Graph Sequence Neural Networks modify GNNs
to use gated recurrent units and to output sequences (Li et al., 2015).

One back-propagation per output: each node embedding is an output.

Reviewer 1: rotational invariance.

Your analysis is correct, the proposed filters are isotropic, i.e. they are
oblivious to the orientation of an edge. The learned filters therefore provide
rotational invariance. Whether this is a limitation or an asset depends on the
problem and should be verified empirically, as for any invariance (translation,
scale). In an experiment (not in the paper) where we did randomly rotate (from
-90° to 90°) each MNIST digit (no data augmentation), we obtained 91.47%
accuracy while classical CNNs got 70.65%. Rotational invariance can therefore be
an asset. Moreover, data augmentation schemes have used rotated versions of
(part of) images to force a network to learn this invariance, and Spatial
Transformer Networks [1], who learn invariance to translation, scale, rotation
and more generic warping, did show state-of-the-art performance on several
benchmarks. We did not include these preliminary results in the paper as we want
to test on larger image datasets the impact of this invariance. As part of this
future work, we'll investigate the possibility to build our filters on
anisotropic Laplacians [2], which should allow us to learn directional /
oriented filters.

Reviewer 3.

The limited results in Table 4 could be due to the small size (~19k documents)
of the 20NEWS dataset. We performed the same task on RCV1 (800k documents), and
obtained...
Unfortunately after the submission deadline. Or say that Bruna did it.

Reviewer 6.

Thanks for your encouragement. Please note that most of the recent works are
based on a spatial formulation. What are the limitations? We look forward to
compare all these approaches.

Reviewer 7: single graph.

As you rightfully pointed out, our spectral construction does not transfer
directly to other graphs. That is an advantage of spatial constructions?. We do
however believe that graphs from a certain class (e.g. social graphs) share a
similar spectrum, which would allow to learn and transfer the Chebyshev
coefficients across graphs. That is a planned future work.?

Reviewer 7: pooling on graphs.

A pooling strategy on arbitrary graphs will always? require some sort of
coarsening, i.e. a way to group similar vertices. As we pointed out, there
exists plenty of methods for multi-level clustering. The proposed CNN on graphs
is agnostic to the choice of a particular algorithm, and we hope that future
research in the area will bring better algorithms. Moreover, the coarsening step
only has to be performed once, before training.

Reviewer 7: MNIST experiment.

You are right, we did experiment with a regular grid on MNIST because it is the
sole way to compare with classical CNNs. The ~3% performance loss you mention
(from Table 3 I guess?) is caused by the use of random graph, i.e. random
connections between pixels. It is clear that in this setting we expect to loose
much performance, which indeed shows that the grid structure is important for
images. We could also build a graph from the pixel values only, which would
approach a regular grid?. Performances are obviously not state-of-the-art, we
only wanted to compare with the classical LeNet-5 and not bother with model
tuning.

Reviewer 7 & 8: 20NEWS experiment, usefulness of the graph.

We did compare with baselines only. The small performance improvement is
probably because the constructed graph is not good enough. There is probably a
better way. Constructing graphs from data is an open research question. This is
however a non-issue when the data naturally lies on a graph structure. Then why
didn't we do it ??

The usefulness of the graph structure obviously depends on the quality of the
graph. That is the point we tried to make in section 6.2: if a given graph does
not improve much on a random graph, then this graph is not well suited for the
problem at hand. If the graph is well constructed, e.g. the grid for MNIST, then
results are good. Moreover, the very recent works in this area [refs] suggest
that this is a research direction of interest.

Reviewer 8 & 9: understanding.

As you pointed out, you do not seem to be familiar with the field of Graph
Signal Processing. Due to space constraints, we had to write densely and
specifically pointed out a review of the field [29]. Moreover, most of the
reviewers seem to have understood the paper well. We did however clean the paper
up and hope it is easier to read.

Reviewer 8: performance.

Our method is both faster and more effective, see Table 6 (Spline filters), than
[13]. The theoretical difference is that our filters are exactly localized in K
hops.

Reviewer 9: natural graphs.

The problem with signals on natural graphs is that, at least for now, not many
datasets are public, and when they are, they are quite small. ?? We do however
plan to make experiments on Wikipedia, using hyperlinks and user activity to
build a graph. This will however require us to build a new dataset.

[1] Spatial Transformer Networks 
[2] anisotropic 
[3] Rotation-invariant convolutional neural networks for galaxy morphology prediction

### 2016-10-29 nips2016/nips2016.tex
large edition for nips camera-ready, first submission (2016-10-27)

\pdfoutput=1

@@

%\usepackage{nips_2016}
%\usepackage[final]{nips_2016}
\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry} % arxiv

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[english]{babel}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{amssymb}
\usepackage{xfrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
%\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{multirow}
%\usepackage{svg}
\usepackage{setspace}
\usepackage{color}

\usepackage{amsmath}

% Use less space
%\usepackage{titlesec}		% space before / after headings
%\titlespacing\section{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
%\usepackage{parskip} \setlength{\parskip}{5pt}
%\setlength{\bibsep}{0.2pt} %comment for arxiv

\setlength{\textfloatsep}{20.0pt plus 2.0pt minus 2.0pt}
\setlength{\floatsep}{20.0pt plus 2.0pt minus 2.0pt}
\setlength{\intextsep}{20.0pt plus 2.0pt minus 2.0pt}

@@

%\author{
%  Michaël Defferrard \\ %\thanks{further info} \\
%  %Michael Defferrard \\ %\thanks{further info} \\
%  %Department of Electrical Engineering \\
%  %École Polytechnique Fédérale de Lausanne (EPFL) \\
%  EPFL, Lausanne, Switzerland \\
%  \texttt{michael.defferrard@epfl.ch} \\
%  \And % \AND
%  Xavier Bresson \\
%  EPFL, Lausanne, Switzerland \\
%  \texttt{xavier.bresson@epfl.ch} \\
%  \And % \AND
%  Pierre Vandergheynst \\
%  EPFL, Lausanne, Switzerland \\
%  \texttt{pierre.vandergheynst@epfl.ch} \\
%}

%%Arxiv
\author{
  Micha\"{e}l Defferrard\hspace{1cm}Xavier Bresson\hspace{1cm} Pierre Vandergheynst\\ \\
  Department of Electrical Engineering \\
  EPFL, Lausanne, Switzerland \\
  {\small \texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} } \\\\
}

\usepackage[acronym]{glossaries}
\newacronym{CNN}{CNN}{Convolutional Neural Network}
\newacronym{SVD}{SVD}{Singular Value Decomposition}
\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
\newacronym{PSD}{PSD}{positive semidefinite}

\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\spn}{span}
\newcommand{\G}{\mathcal{G}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\bO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\eqnref}[1]{(\ref{eq:#1})}
\newcommand{\bruna}{art:BrunaZarembaSzlamLeCun13DLgraphs,
art:HenaffBrunaLeCun15DLgraphs}

@@

Convolutional neural networks (CNNs) have greatly improved state-of-the-art performances in a number of fields, notably computer vision and natural language processing. In this work, we are interested in generalizing the formulation of CNNs from low-dimensional regular Euclidean domains, where images (2D), videos (3D) and audios (1D) are represented, to high-dimensional irregular domains such as social networks or biological networks represented by graphs. This paper introduces a formulation of CNNs on graphs in the context of spectral graph theory. We borrow the fundamental tools from the emerging field of signal processing on graphs, which provides the necessary mathematical background and efficient numerical schemes to design localized graph filters efficient to learn and evaluate.  As a matter of fact, we introduce the first technique that offers the same computational complexity than standard CNNs, while being universal to any graph structure. Numerical experiments on MNIST and 20NEWS demonstrate the ability of this novel deep learning system to learn local, stationary, and compositional features on graphs, as long as the graph is well-constructed.

@@

	%Graphs can either be the natural domain of the data, such as social networks,
	%biological graphs like gene regulatory and brain connectivity networks,
	%telecommunication networks, or constructed from the data, such as similarity
	%graphs.

@@

%Previous works have been developed along this today essential line of research, but they lack of mathematical foundations and efficient learning of filters.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{images/illustrationCNNgraphs}
\caption{Architecture of the proposed CNNs on graphs. Notation: $l$ is the
coarsening level, $x^l$ are the downsampled signals at layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the coarsening exponent, $n_c$ is the number of classes, $y$ is the output signal, and $\theta^l$ is the number of parameters to learn at $l$.}
\label{fig_illustration}
\end{figure}

@@

Convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST}
offer an efficient architecture to extract highly meaningful statistical
patterns in large-scale and high-dimensional datasets. The key success of CNNs
is its advanced ability to learn local stationary structures and compose them to
form multi-scale hierarchical patterns. Precisely, CNNs extract the local
stationarity property of the input data or signals by revealing local features
that are shared across the data domain. These similar features are identified
with localized convolutional filters or kernels, which are learned from the
data. Convolutional filters are shift- or translation-invariant filters, meaning
they are able to recognize identical features independently of their spatial
locations. Localized kernels or compactly supported filters refer to filters
that extract local features independently of the input data size, with a support
size that can be much smaller than the input size.

%\newpage
The compositional local stationarity property of CNNs can be efficiently
implemented on regular grids, like image and sound domains, as convolutional, 
downsampling and pooling operators are mathematically well-defined on such discretized
Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on
low-dimensional regular lattices. But not all data lie on regular grids. User
data on social networks, gene data on biological regulatory networks, log data
on telecommunication networks, or text documents on words' embedding are
important examples of data lying on irregular or non-Euclidean domains.
Unfortunately, CNNs cannot be directly applied to such complex domains.% as the building blocks of CNNs, convolution, downsampling, and pooling, are only defined for regular grids.

%A generalization to irregular grids is not straightforward as the building blocks of CNNs, convolution, downsampling, and pooling, are only defined for regular grids.

A generalization to irregular grids is not straightforward as the standard operators of convolution, downsampling, and pooling, are only defined for regular grids. This makes this extension challenging, both theoretically and implementation-wise. This work leverages the recent works of
\cite{pro:GregorLeCun10LRF, pro:CoatesNg11LRF,
art:BrunaZarembaSzlamLeCun13DLgraphs, art:HenaffBrunaLeCun15DLgraphs,
art:HammondVandergheynstGribonval11GraphWav,
art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to introduce an
efficient implementation of CNNs on irregular domains represented here as
graphs. Graphs are universal representations of heterogeneous pairwise
relationships between possibly high-dimensional data. Graphs can encode complex
geometric structures, and can be studied with strong mathematical tools such as
spectral graph theory \cite{book:Chung97Spectral}, a blend between graph theory and harmonic analysis.

@@

The major bottleneck of generalizing CNNs to graphs, and one of the primary goals of
this work, is the definition of graph localized filters which are efficient to
evaluate and learn.
%that are the core elements to extract local stationarity structures.
We will make use of recent tools developed in the context of graph signal
processing (GSP) \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG} to achieve such goals. Precisely, the main contributions of this work are summarized below:
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A graph spectral theoretical formulation of CNNs on graphs built on
	established tools in GSP.
\item \textbf{Strictly localized filters.} Enhancing \cite{\bruna}, the proposed spectral
	filters are provable to be strictly localized in a ball of radius $K$, i.e. $K$ hops from
	the central vertex.  
\item \textbf{Low computational complexity.} The evaluation complexity of our filters is
	linear w.r.t. the filters support's size $K$ and the number of edges $|\E|$. Importantly, as most real-world graphs are highly sparse, we have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor (NN) graphs, leading to a linear complexity w.r.t the input data size. Moreover,
	this method avoids the Fourier basis altogether, thus the expensive
	eigenvalue decomposition (EVD) necessary to compute it as well as the need
	to store the basis, a matrix of size $n^2$. That is especially relevant when
	working with limited GPU memory. Besides the data, our method only requires
	to store the Laplacian, a sparse matrix of $|\E|$ non-zero values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices as a binary tree structure, is analog to pooling of
	1D signals.
\item \textbf{Experimental results.} We present many experiments that ultimately
	show that our formulation is (1) a useful model, (2) computationally
	efficient and (3) superior both in accuracy and complexity to the pioneer spectral graph CNNs introduced in \cite{\bruna}. We also show that
	our graph formulation performs similarly as a classical CNNs on MNIST and study the
	impact of various graph constructions on classification performance.
\end{enumerate}

\noindent
Eventually, generic and efficient implementations of CNNs on
graphs, as proposed here and related works \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,pro:MasciBoscainiBronsteinVandergheynst15GeoDL,
art:MasciBoscainiBronsteinVandergheynst15ShapeNet,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs}, is essential to pave the way to the development of a new class of deep learning techniques for graph-based data.

\section{Related Works}
\subsection{Graph Signal Processing} GSP is an emerging field that aims at
bridging the gap between signal processing techniques like wavelet analysis \cite{book:Mallat99wavelets} and
graph theory such as spectral graph analysis
\cite{art:BelkinNiyogi05LaplaBeltrami, art:VonLuxburg07Tutorial}. A goal is to
generalize fundamental analysis operations for signals from regular grids to
irregular structures embodied by graphs. We refer the reader to
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}
 for an
introduction of the field. Standard operations on grids such as convolution,
translation, filtering, dilatation, modulation or downsampling do not extend
directly to graphs and thus require new mathematical definitions while keeping
the original intuitive concepts. In this context, the authors of
\cite{art:HammondVandergheynstGribonval11GraphWav, art:CoifmanLafon06DifMap,
pro:GavishNadlerCoifman10GraphHaar} revisited the construction of wavelet
operators on graphs. In \cite{art:ShumanFarajiVandergheynst16PyramTrans,
art:RamEladCohen11TreeWavelets}, the authors designed a technique to perform
mutli-scale pyramid transforms on graphs. The works of
\cite{pro:TsitsveroBarbarossa15Uncert, pro:PasdeloupAlamiGriponRabbat15Uncert,
art:PerraudinRicaudShumanVandergheynst16Uncert} redefined uncertainty principles
on graphs, and showed that intuitive concepts may be lost, but can also produce
enhanced localization principles for signals on graphs. In
\cite{pro:HammondRaoaroorJacquesVandergheynst10LassoGraWav}, it was shown how to
carry out lasso-based signal regularization on graphs, and studied the
intertwined relationships between smoothness and sparsity on graphs. In
\cite{pro:TremblayPuyGribonvalVandergheynst16CompSpecClus}, the authors
investigated compressed sensing recovery conditions for graph spectral signal
analysis.

In the future, we expect to
further benefit from newly developed tools in GSP to enhance our mathematical formulation of
CNNs on graphs. This paper introduces the mathematical and computational foundations of
such models, while future developments may consider to enhance the building
blocks, namely graph filtering and coarsening.

>>>>

@@

%\usepackage{nips_2016}
\usepackage[final]{nips2016}
%\usepackage[top=1.2in, bottom=1.2in, left=1.2in, right=1.2in]{geometry} % arxiv

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[english]{babel}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}       % hyperlinks
%\usepackage{url}            % simple URL typesetting
%\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{xfrac}       % compact symbols for 1/2, etc.
%\usepackage{microtype}      % microtypography
\usepackage{subfigure}
\usepackage{graphicx}
%\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{multirow}
%\usepackage{svg}
\usepackage{setspace}
\usepackage{color}
\usepackage{capt-of}

% Use less space
\usepackage{titlesec}		% space before / after headings
\titlespacing\section{0pt}{6pt plus 1pt minus 2pt}{1pt plus 1pt minus 2pt}
\titlespacing\subsection{0pt}{6pt plus 1pt minus 2pt}{1pt plus 1pt minus 2pt}
%\titlespacing\subsubsection{0pt}{10pt plus 4pt minus 2pt}{0pt plus 2pt minus 2pt}
\setlength{\bibsep}{1.5pt} %comment for arxiv

\usepackage{parskip} \setlength{\parskip}{0pt} \setlength{\parindent}{1em}

\setlength{\textfloatsep}{7.2pt plus 1.0pt minus 2.0pt}
\setlength{\floatsep}{7.2pt plus 1.0pt minus 2.0pt}
\setlength{\intextsep}{7.2pt plus 1.0pt minus 2.0pt}

\setcounter{topnumber}{3}
%\setcounter{bottomnumber}{8}
%\setcounter{totalnumber}{8}

@@

\author{
  Michaël Defferrard \\ %\thanks{further info} \\
  %Department of Electrical Engineering \\
  %École Polytechnique Fédérale de Lausanne (EPFL) \\
  EPFL, Lausanne, Switzerland \\
  \texttt{michael.defferrard@epfl.ch} \\
  \And % \AND
  Xavier Bresson \\
  %School of Computer Science and Engineering \\
  Nanyang Technological University, Singapore \\
  \texttt{xavier.bresson@ntu.edu.sg} \\
  \And % \AND
  Pierre Vandergheynst \\
  %Department of Electrical Engineering \\
  EPFL, Lausanne, Switzerland \\
  \texttt{pierre.vandergheynst@epfl.ch} \\
}

%%Arxiv
%\author{
%  Micha\"{e}l Defferrard\hspace{1cm}Xavier Bresson\hspace{1cm} Pierre Vandergheynst\\ \\
%  Department of Electrical Engineering \\
%  EPFL, Lausanne, Switzerland \\
%  {\small \texttt{\{michael.defferrard,xavier.bresson,pierre.vandergheynst\}@epfl.ch} } \\\\
%}

%\usepackage[acronym]{glossaries}
%\newacronym{CNN}{CNN}{Convolutional Neural Network}
%\newacronym{SVD}{SVD}{Singular Value Decomposition}
%\newacronym{SGD}{SGD}{Stochastic Gradient Descent}
%\newacronym{PSD}{PSD}{positive semidefinite}

\DeclareMathOperator*{\diag}{diag}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\spn}{span}
\newcommand{\G}{\mathcal{G}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\bO}{\mathcal{O}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\eqnref}[1]{(\ref{eq:#1})}

@@

In this work, we are interested in generalizing convolutional neural networks
(CNNs) from low-dimensional regular grids, where image, video and speech are
represented, to high-dimensional irregular domains, such as social networks,
brain connectomes or words’ embedding, represented by graphs.
We present a formulation of CNNs in the context of spectral graph theory, which
provides the necessary mathematical background and efficient numerical schemes
to design fast localized convolutional filters on graphs. Importantly, the
proposed technique offers the same linear computational complexity and constant
learning complexity as classical CNNs, while being universal to any graph
structure. Experiments on MNIST and 20NEWS demonstrate the ability of this
novel deep learning system to learn local, stationary, and compositional
features on graphs.

@@

@@

%Previous works have been developed along this today essential line of
%research, but they lack of mathematical foundations and efficient learning of
%filters.

@@

%Convolutional neural networks (CNNs) have greatly improved state-of-the-art performances in a number of fields, notably computer vision.

	%Graphs can either be the natural domain of the data, such as social networks,
	%biological graphs like gene regulatory and brain connectivity networks,
	%telecommunication networks, or constructed from the data, such as similarity
	%graphs.
% Constructed graphs make sense because of comp complexity, as evidenced by LRF.

Convolutional neural networks \cite{pro:LeCunBottouBengioHaffner98MNIST} offer
an efficient architecture to extract highly meaningful statistical patterns in
large-scale and high-dimensional datasets. The ability of CNNs to learn local
stationary structures and compose them to form multi-scale hierarchical
patterns has led to breakthroughs in image, video, and sound recognition tasks
\cite{art:LeCunBengioHinton15DL}. Precisely, CNNs extract the local
stationarity property of the input data or signals by revealing local features
that are shared across the data domain. These similar features are identified
with localized convolutional filters or kernels, which are learned from the
data. Convolutional filters are shift- or translation-invariant filters,
meaning they are able to recognize identical features independently of their
spatial locations. Localized kernels or compactly supported filters refer to
filters that extract local features independently of the input data size, with
a support size that can be much smaller than the input size.

\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{images/illustrationCNNgraphs}
\caption{Architecture of the proposed CNN on graphs.
	%Notation: $l$ is the coarsening level, $x^l$ are the downsampled signals at
	%layer $l$, $\G^l$ is the coarser graph, $g_{\theta^{K_l}}$ are the spectral
	%filters at layer $l$, $x_g^l$ are the filtered signals, $p_l$ is the
	%coarsening exponent, $n_c$ is the number of classes, $y$ is the output
	%signal, and $\theta^l$ is the number of parameters to learn at $l$.
}
\label{fig_illustration}
\end{figure}

% Eventually, generic and efficient implementations of CNNs on graphs, as
% proposed here and related works
% \cite{pro:GregorLeCun10LRF,pro:CoatesNg11LRF,masci2015geodesic,art:BrunaZarembaSzlamLeCun13DLgraphs,art:HenaffBrunaLeCun15DLgraphs},
% is essential to pave the way to the development of a new class of deep learning
% techniques for graph-based data.

%The compositional local stationarity property of CNNs can be efficiently
%implemented on regular grids, like image and sound domains, as convolutional, 
%downsampling and pooling operators are mathematically well-defined on such discretized
%Euclidean spaces. This has led to the breakthrough of CNNs in image, video, and
%sound recognition tasks \cite{art:LeCunBengioHinton15DL} as these data lie on
%low-dimensional regular lattices. But not all data lie on regular grids. User
%data on social networks, gene data on biological regulatory networks, log data
%on telecommunication networks, or text documents on words' embedding are
%important examples of data lying on irregular or non-Euclidean domains.
%Unfortunately, CNNs cannot be directly applied to such complex domains.

@@

%Unfortunately, CNNs cannot be directly applied to such complex domains.

User data on social networks, gene data on biological regulatory networks, log
data on telecommunication networks, or text documents on word embeddings are
important examples of data lying on irregular or non-Euclidean domains which
can be structured with graphs, which are universal representations of
heterogeneous pairwise relationships. Graphs can encode complex geometric
structures and can be studied with strong mathematical tools such as spectral
graph theory \cite{book:Chung97Spectral}.

A generalization of CNNs to graphs is not straightforward as the convolution
and pooling operators are only defined for regular grids. This makes this
extension challenging, both theoretically and implementation-wise. The major
bottleneck of generalizing CNNs to graphs, and one of the primary goals of this
work, is the definition of localized graph filters which are efficient to
evaluate and learn.
Precisely, the main contributions of this work are summarized below.
\begin{enumerate}
%\setlength{\itemsep}{0pt} \setlength{\parskip}{0pt}
\item \textbf{Spectral formulation.} A spectral graph theoretical formulation
	of CNNs on graphs built on established tools in GSP
	\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
\item \textbf{Strictly localized filters.} Enhancing
	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}, the proposed spectral filters
	are provable to be strictly localized in a ball of radius $K$, i.e. $K$
	hops from the central vertex.  
\item \textbf{Low computational complexity.} The evaluation complexity of our
	filters is linear w.r.t. the filters support's size $K$ and the number of
	edges $|\E|$. Importantly, as most real-world graphs are highly sparse, we
	have $|\E| \ll n^2$ and $|\E| = kn$ for the widespread $k$-nearest neighbor
	(NN) graphs, leading to a linear complexity w.r.t the input data size.
	Moreover, this method avoids the Fourier basis altogether, thus the
	expensive eigenvalue decomposition (EVD) necessary to compute it as well as
	the need to store the basis, a matrix of size $n^2$. That is especially
	relevant when working with limited GPU memory. Besides the data, our method
	only requires to store the Laplacian, a sparse matrix of $|\E|$ non-zero
	values.
\item \textbf{Efficient pooling.} We propose an efficient pooling strategy on
	graphs which, after a rearrangement of the vertices as a binary tree
	structure, is analog to pooling of 1D signals.
\item \textbf{Experimental results.} We present multiple experiments that
	ultimately show that our formulation is (i) a useful model, (ii)
	computationally efficient and (iii) superior both in accuracy and
	complexity to the pioneer spectral graph CNN introduced in
	\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. We also show that our graph
	formulation performs similarly as a classical CNNs on MNIST and study the
	impact of various graph constructions on classification performance. The
	TensorFlow \cite{abadi_tensorflow_2016} code to reproduce our results and
	apply the model to other data is available as an open-source
	software.\footnote{\url{https://github.com/mdeff/cnn_graph}}
\end{enumerate}

\section{Proposed Technique}

Generalizing CNNs to graphs requires three fundamental steps: (i) the design of
localized convolutional filters on graphs, (ii) a graph coarsening procedure
that groups together similar vertices and (iii) a graph pooling operation that
trades spatial resolution for higher filter resolution.

\subsection{Learning Fast Localized Spectral Filters} \label{sec:filters}

There are two strategies to define convolutional filters; either from a spatial
approach or from a spectral approach. By construction, spatial approaches
provide filter localization via the finite size of the kernel. However,
although graph convolution directly in the spatial domain is conceivable, it
also faces the challenge of matching local neighborhoods, as pointed out in
\cite{art:BrunaZarembaSzlamLeCun13DLgraphs}. Consequently, there is no unique
mathematical definition of translation on graphs from a spatial perspective. On
the other side, a spectral approach provides a well-defined translation
operator on graphs via convolutions with a Kronecker delta implemented in the
spectral domain \cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}.
The convolution theorem \cite{book:Mallat99wavelets} defines convolutions as
linear operators that diagonalize in the Fourier basis (represented by the
eigenvectors of the Laplacian operator). However, a filter defined in the
spectral domain is not naturally localized and translations are costly due to
the $\bO(n^2)$ multiplication with the graph Fourier basis. Both limitations
can however be overcome with a special choice of filter parametrization.

\paragraph{Graph Fourier Transform.} We are interested in processing signals
defined on undirected and connected graphs $\G=(\V,\E,W)$, where $\V$ is a
finite set of vertices with $|\V|=n$, $\E$ is a set of edges and $W$ is a
weighted adjacency matrix encoding the connection weight between two vertices. A
signal $x: \V \rightarrow \R$ defined on the nodes of the graph may be regarded
as a vector $x \in \R^n$ where $x_i$ is the value of $x$ at the $i^{th}$ node.
An essential operator in spectral graph analysis is the graph Laplacian
\cite{book:Chung97Spectral}, which combinatorial definition is $L = D - W \in
\R^{n \times n}$ where $D \in \R^{n \times n}$ is the diagonal degree matrix
with $D_{ii} = \sum_j W_{ij}$, and normalized definition is $L = I_n -
D^{-1/2} W D^{-1/2}$ where $I_n$ is the identity matrix. As $L$ is a real
symmetric positive semidefinite matrix, it has a complete set of orthonormal
eigenvectors $\{u_l\}_{l=0}^{n-1} \in \R^n$, known as the graph Fourier modes,
and their associated ordered real nonnegative eigenvalues
$\{\lambda_l\}_{l=0}^{n-1}$, identified as the frequencies of the graph. The
Laplacian is indeed diagonalized by the Fourier basis $U=[u_0, \ldots, u_{n-1}]
\in \R^{n \times n}$ such that $L = U \Lambda U^T$ where $\Lambda =
\diag([\lambda_0, \ldots, \lambda_{n-1}]) \in \R^{n \times n}$. We can now
define the graph Fourier transform (GFT) of a spatial signal $x \in \R^n$ as
$\hat{x} = U^T x \in \R^n$, and its inverse as $x = U \hat{x}$
\cite{art:ShumanNarangFrossardOrtegaVandergheynst13ReviewSPG}. Similarly to the
Fourier transform on Euclidean spaces, the GFT enables the formulation of
fundamental operations such as filtering.

\paragraph{Polynomial parametrization for localized filters.} There are however
two limitations with non-parametric filters: (i) they are not localized in
space and (ii) their learning complexity is in $\bO(n)$, the dimensionality of
the data. These issues can be overcome with the use of a polynomial filter
% a Laplacian-based polynomial spectral filter
\begin{equation} \label{eq:filt_poly}
	g_\theta(\Lambda) = \sum_{k=0}^{K-1} \theta_k \Lambda^k,
\end{equation}
where the parameter $\theta \in \R^K$ is a vector of polynomial coefficients.
The value at vertex $j$ of the filter $g_\theta$ centered at vertex $i$ is
given by $(g_\theta(L) \delta_i)_j = (g_\theta(L))_{i,j} = \sum_k \theta_k
(L^k)_{i,j}$, where the kernel is translated via a convolution with a Kronecker
delta function $\delta_i \in \R^n$. By \cite[Lemma
5.2]{art:HammondVandergheynstGribonval11GraphWav}, $d_\G(i,j) > K$ implies
$(L^K)_{i,j} = 0$, where $d_\G$ is the shortest path distance, i.e. the minimum
number of edges connecting two vertices on the graph.  Consequently, spectral
filters represented by $K^\text{th}$-order polynomials of the Laplacian are
exactly $K$-localized. Besides, their learning complexity is $\bO(K)$, the
support size of the filter, and thus the same complexity as classical CNNs.

% While there is many possible choice of recursive polynomials
\paragraph{Recursive formulation for fast filtering.} While we have shown how
to learn localized filters with $K$ parameters, the cost to filter a signal $x$
as $y = Ug_\theta(\Lambda)U^Tx$ is still high with $\bO(n^2)$ operations
because of the multiplication with the Fourier basis $U$. A solution to this
problem is to parametrize $g_\theta(L)$ as a polynomial function that can be
computed recursively from $L$, as $K$ multiplications by a sparse $L$ costs
$\bO(K|\E|) \ll \bO(n^2)$. One such polynomial, traditionally used in GSP to
approximate kernels (like wavelets), is the Chebyshev expansion
\cite{art:HammondVandergheynstGribonval11GraphWav}. Another option, the Lanczos
algorithm \cite{art:SusnjaraPerraudinKressnerVandergheynst15Lanczos}, which
constructs an orthonormal basis of the Krylov subspace $\mathcal{K}_K(L,x) =
\spn\{x,Lx,\ldots,L^{K-1}x\}$, seems attractive because of the coefficients'
independence. It is however more convoluted and thus left as a future work.
